---
title: Virtual Buoy Example 
author: NOAA CoastWatch
date: August 2023
output:
  md_document
---


> Updated August 2023 <br/>

There are buoys in many locations around the world that provide data streams of oceanic and atmospheric parameters. The data are often available through data centers like the those National Data Buoy Center (NDBC https://www.ndbc.noaa.gov) and ARGO floats program (http://www.argo.ucsd.edu). In situ buoy data are widely used to monitor environmental conditions.

In the absence of in situ buoy data, whether the buoy operation is discontinued, interrupted, or limited, satellite data with close temporal and spatial coverage can be used to create a dataset in the format of in situ buoy data.

## Objective

This tutorial will demonstration how to transform satellite data to a dataset in an in-situ data format. 

## The tutorial demonstrates the following techniques

-   Downloading and saving a netcdf file from CoastWatch ERDDAP data server
-   Accessing satellite data and metadata 
-   Resampling satellite data into a lower resolution timestep
-   Data-wranging to transform the satellite data into in situ buoy data format
-   Validating the transformed data with the actual buoy data
-   Running a linear regression model and generating statistics
-   Plotting time series data

## Datasets used

-  <a href="https://coastwatch.pfeg.noaa.gov/erddap/griddap/nesdisBLENDEDsstDNDaily.graph">The sea surface temperature (SST) satellite data</a> from NOAA Geo-polar blended analysis are used for transforming to buoy data format

-  <a href="https://coastwatch.pfeg.noaa.gov/erddap/tabledap/cwwcNDBCMet.graph?time%2Cwtmp%2Cwd&station=%2246259%22&time%3E=2020-09-15T00%3A00%3A00Z&time%3C=2022-09-15T00%3A00%3A00Z&.draw=markers&.marker=5%7C5&.color=0x000000&.colorBar=%7C%7C%7C%7C%7C&.bgColor=0xffccccff"> NDBC Standard Meteorological Buoy Data </a> was used for validating or ground truthing

## References

- <a href="https://coastwatch.pfeg.noaa.gov/data.html">NOAA CoastWatch Westcoast Node Data Catalog</a>
- <a href="https://www.ndbc.noaa.gov/download_data.php?filename=46259h2017.txt.gz&dir=data/historical/stdme ">NOAA National Data Buoy Center</a>

## Install required packages and load libraries
```{r install,message=FALSE,warning=FALSE}

# Function to check if pkgs are installed, and install any missing pkgs
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,dep=TRUE,repos='http://cran.us.r-project.org')
    if(!require(x,character.only = TRUE)) stop(x, " :Package not found")
  }
}


# Create list of required packages
list.of.packages <- c("rerddap", "plotdap", "parsedate", "ggplot2", "rerddapXtracto",
                       "maps", "mapdata", "RColorBrewer")

# Create list of installed packages
pkges = installed.packages()[,"Package"]

# Install and load all required pkgs
for (pk in list.of.packages) {
  pkgTest(pk)
}

```

## Download the Sea Surface Temperature (SST) Data

The Sea Surface Temperature (SST) is the NOAA GeoPolar Blended SST dataset (in Celcius) from many satellite sensors to obtain good daily coverage of the globe at 5km resolution, and then an interpollation method is applied to fill in data gaps. 

**Data**

*Sea-Surface Temperature, NOAA Geo-polar Blended Analysis Day+Night, GHRSST,
Near Real-Time, Global 5km, 2019-Present, Daily*



## Sending data request to CoastWatch ERDDAP Server
```{r}
library('utils')

# Set ERDDAP URL 
url="https://coastwatch.pfeg.noaa.gov/erddap/griddap/nesdisBLENDEDsstDNDaily.csv?analysed_sst%5B(2020-01-01T12:00:00Z):1:(last)%5D%5B(34.775):1:(34.775)%5D%5B(-121.475):1:(-121.475)]"


# Set file name
fname = 'sst.csv'

# Download file
download.file(url, fname)

# Read into data frame, skip the first row
sst_df <- read.csv(fname, skip=1, header=TRUE)

# Show the first 3 rows
head(sst_df, 3)
```

```{r erddap}

# Deleting all cached files
rerddap::cache_delete_all(force = TRUE)

# Set dataset ID
dataset <- 'esa-cci-chla-monthly-v6-0'

# Get data information from ERDDAP server
dataInfo <- rerddap::info(dataset, url= "https://oceanwatch.pifsc.noaa.gov/erddap")

# Display the metadata
dataInfo
```

