---
title: "Transforming coordinates from crs to crs"
output: html_document
---

> history \| Create July 2023 \| Updated August 2023<br/>


<a href="https://polarwatch.noaa.gov/" target="_blank">NOAA PolarWatch</a> distributes gridded and tabular oceanographic data for polar regions.  Satellite data include geospatial information and most of them are in geographical coordinates (latitude and longitude).  Since the spatial coverage of many PolarWatch satellite data products, are within polar regions, the geospatial information is often represented in polar projected coordinates. (<a href="https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection" target="_blank">Polar Stereographic Projections</a>).  

When working with multiple satellite data sets from different crs coordinates, it makes it challenging to do data analysis and the coordinate transformation may be needed. 

In this exercise, we will learn to transform coordinates of one projection to another.

**The exercise demonstrates the following techniques:**
-   Downloading and saving a netcdf file from PolarWatch ERDDAP data server
-   Accessing satellite data and metadata in polar stereographic projection 
-   Transforming coordinates using EPSG codes
-   Mapping data using the transformed coordinates

**Dataset:**
    
   *Sea Ice Concentration*, NOAA/NSIDC Climate Data Record V4, Southern Hemisphere, 25km, Science Quality, 1978-2022, Monthly

**R Packages**

-   ncdf4 (reading data and metadata in netCDF format)
-   ggplot2, RColorBrewer, scales (mapping)
-   reshape2 (data manipulation)
-   rgdal, sf (projection)

### Install required packages
```{r setup_packages, message=FALSE, warning=FALSE}

# Function to check if pkgs are installed, install missing pkgs, and load
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,dep=TRUE,repos='http://cran.us.r-project.org')
    if(!require(x,character.only = TRUE)) stop(x, " :Package not found")
  }
}

list.of.packages <- c("raster", "ncdf4")

# create list of installed packages
pkges = installed.packages()[,"Package"]
for (pk in list.of.packages) {
  pkgTest(pk)
}

```

## Antarctic Sea Ice Concentration Data

The sea ice concentration (SIC) data are percent (proportion) coverage of ice within each grid cell.  The SIC dataset used in this exercise is produced by the NOAA/NSIDC from passive microwave sensors using the Climate Data Record (CDR).  The data are projected in the southern polarstereographic projection (**EPSG:3031**). This sea ice concentration data values in the dataset are represented as proportions ranging from 0 to 1. The resolution is 25km, meaning each grid in this data set represents a value that covers a 25km by 25km area.  Please note that other values beyond 1 represents fill values.  More detailed information can be found in the NSIDC Data Product User Guide (linke below)

**Userful links**

-   <a href="https://polarwatch.noaa.gov/catalog/ice-sq-nh-nsidc-cdr-v4/preview/?dataset=daily&var=cdr_seaice_conc&time_min=2022-05-31T00:00:00Z&time_max=2022-05-31T00:00:00Z&proj=epsg3413&colorBar=KT_ice|||0|1|" target="_blank">NOAA PolarWatch Data Product Page (download, preview)</a>
-   <a href="https://nsidc.org/data/g02202/versions/4" target="_blank">NSIDC Data Product Description</a>
-   <a href="https://nsidc.org/sites/default/files/g02202-v004-userguide_1_1.pdf" target="_blank">NSIDC Data Product User Guide (pdf)</a>


```{r}
# download the sea ice data NetCDF file
url <- "https://polarwatch.noaa.gov/erddap/griddap/nsidcG02202v4shmday.nc?cdr_seaice_conc_monthly[(2022-12-01T00:00:00Z):1:(2022-12-01T00:00:00Z)][(4350000.0):1:(-3950000.0)][(-3950000.0):1:(3950000.0)]"

sic <- download.file(url, destfile="sic.nc", mode='wb')

# file open
ds <- nc_open('sic.nc')

# print metadata
#print(ds)

# get data into r variables 
xgrid <- ncvar_get(ds, "xgrid")
ygrid <- ncvar_get(ds, "ygrid")
sic <- ncvar_get(ds, "cdr_seaice_conc_monthly")  #lat and lon

# close 
nc_close(ds)
```

## Transform data to a new CRS
```{r}

# create raster
ext <- extent(min(xgrid), max(xgrid), min(ygrid), max(ygrid))
rastered_sic <- raster(t(sic)) 
extent(rastered_sic)<- ext
names(rastered_sic) <- 'sic'
# set the crs of the data to 3031 (polar stereographic)
crs(rastered_sic) <- CRS("+init=epsg:3031")

# transform to EPSG 4326
# note: supressWarnings applied as the function throws warning for invalid latitude and longitude
sic_to_4326 <- suppressWarnings(projectRaster(rastered_sic, crs = "+init=epsg:4326"))

```


## Map data with new coordinates on a global map
```{r}
# transform raster to data frame
df <- as.data.frame(sic_to_4326, xy=TRUE) 

View(df)
# get world map
world_map <- map_data("world")


ggplot() +
  geom_raster(data = df, aes(x = x, y = y, fill = layer)) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), 
               fill = NA, color = "black") +
  coord_sf(xlim = c(-180, 180), ylim = c(-90, 90)) +
  theme_minimal() +
  labs(fill = "Sea ice concentration")
ggplot(data = df, aes(x = xgrid, y = ygrid) ) + 
        geom_tile(aes(fill=sic)) + 
       coord_fixed(ratio = 1) + 
       scale_y_continuous(labels = comma) + 
       scale_x_continuous(labels = comma) +
       scale_fill_gradientn(colours=rev(brewer.pal(n = 3, name = "Blues")),na.value="tan")+
       ggtitle("SIC with polar bear tracks on Polar (red) Steregraphic projection")+
      geom_point(data=df, aes(x=x, y=y), color="red", size=0.5)
     

        
```


